---
title: "W5_NYPD analysis"
author: "Naomichi Kadota"
date: "2023/11/18"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 0) load the library

```{r library, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(lubridate)
library(patchwork)
```

# 1) Read the CSV file
(COVID-19, by JHU_CSSE)

```{r read_csv_file, include=TRUE}
url_in <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/"
file_names <- c("time_series_covid19_confirmed_global.csv",
                "time_series_covid19_deaths_global.csv",
                "time_series_covid19_confirmed_US.csv",
                "time_series_covid19_deaths_US.csv")
urls <- str_c(url_in, file_names)

global_cases <- read.csv(urls[1])
global_deaths <- read.csv(urls[2])
#US_cases <- read.csv(urls[3])
#US_deaths <- read.csv(urls[4])
```

# 2) Data manipulation
Let's follow the manipulation demonstrated in the video of week-3, and then extract and merge the data of cases and deaths in Japan (out of my interest).

```{r global > JP > merge case & death}
## simplify global data
global_cases <- global_cases %>% 
  pivot_longer(cols = -c('Province.State',
                         'Country.Region',Lat,Long),
               names_to = "date",
               values_to = "case") %>% 
  select(-c(Province.State, Lat, Long))
global_deaths <- global_deaths %>% 
  pivot_longer(cols = -c('Province.State',
                         'Country.Region',Lat,Long),
               names_to = "date",
               values_to = "death") %>% 
  select(-c(Province.State, Lat, Long))

## extract JP data; 1143obs(20/1/12~23/3/09)
JP_cases<- global_cases %>%
  filter(Country.Region=="Japan") %>%
  select(-c(Country.Region))
JP_deaths<- global_deaths %>%
  filter(Country.Region=="Japan") %>% 
  select(-c(Country.Region))
## merge
JP_data <- left_join(JP_cases, JP_deaths, by = "date")
JP_data <- JP_data %>%
  mutate(date = sub("X", "", date),  # Remove the 'X' prefix
         date = mdy(date))

```

# 3) Overview
```{r overview}
summary(JP_data)
P1.1 <- ggplot(JP_data, aes(x = date, y = case))+
  geom_line(colour="blue")
P1.2 <- ggplot(JP_data, aes(x = date, y = death))+
  geom_line(colour="red")
P1.3 <- ggplot(JP_data, aes(x = date))+
  geom_line(aes(y = case, color = "Cases")) +
  geom_line(aes(y = death, color = "Deaths")) +
  scale_color_manual(values = c("Cases" = "blue", "Deaths" = "red")) +
  labs(color = "Variable")
P1.1+P1.2 + plot_layout(ncol = 1)
P1.3

```
 - The plots display the total number of COVID-19 cases and deaths reported.
 - It's noteworthy that the number of deaths is significantly lower than the total number of cases.
 - I hadn't realized the scale difference was so substantial; at first, I even suspected a coding error on my part. However, the plots have provided me with a clearer insight.

# 4) Analyze time series data (taking lag)

```{r lag}
JP_data <- JP_data %>% 
  mutate(new_case = case - lag(case),
         new_death = death - lag(death))

P2.1 <- ggplot(JP_data, aes(x = date, y = new_case))+
  geom_line(colour="blue")
P2.2 <- ggplot(JP_data, aes(x = date, y = new_death))+
  geom_line(colour="red")
P2.1+P2.2 + plot_layout(ncol = 1)

```
 - The patterns of reported cases and deaths are roughly the same.
 - It might be noteworthy that the proportion of deaths prior to 2022 appears relatively high, especially when considering the fewer reported cases compared to the numbers after 2022.

# 5) Modeling
Can we predict the number of new_death from the number of new_case?
```{r simple liner modeling}
JP_data2 <- na.omit(JP_data)
lmod <- lm(new_death~new_case, JP_data2)
summary(lmod)
ggplot(JP_data2, aes(x=new_case, y=new_death))+
  geom_point()

```
According to the regression analysis, the coefficient appears to have a significant impact. However, given the unevenness in the data distribution, the soundness of this interpretation may be questionable.

# 6) Discussion of Potential Bias
 - This analysis focuses on cases in Japan, so the findings may not be generalizable to other countries.
 - As a principle, the case count may include multiple reports for the same individuals, whereas the number of deaths is counted only once per individual. This difference could cause some distortion when interpreting trends.