---
title: "W5_NYPD analysis"
author: "(Assignment)"
date: "2023/11/11"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 0) load the library

```{r library, warning=FALSE}
library(tidyverse)
library(ggplot2)
```

## 1) Read the CSV file (NYDP Shooting Incident) 

```{r read_csv_file, include=TRUE}
NYPD <- read.csv("https://data.cityofnewyork.us/api/views/5ucz-vwe8/rows.csv?accessType=DOWNLOAD")
```

## 2) Overview the data

```{r summary_of_NYPD}
summary(NYPD)
head(NYPD)
str(NYPD)

na_columns <- NYPD %>% 
  select_if(function(x) any(is.na(x))) %>% 
  names()
na_columns # identify which columns contains NA values
```
Some rows lack spacial data, but this time we are to focus on "age" and "sex" of the victim. So there seems to be no such a big problem even if we the whole data.

We have no specific theory to test at the present...
Let's just explore **"VIC_AGE_GROUP"**, *and* **"VIC_SEX"**!

## 3-A) Plot; univariate

```{r barplot age dist, echo=TRUE}
NYPD.2v = NYPD[,c("VIC_AGE_GROUP","VIC_SEX")]
NYPD.2v[] <- lapply(NYPD.2v, factor)
str(NYPD.2v)

table(NYPD.2v$VIC_AGE_GROUP)
barplot(table(NYPD.2v$VIC_AGE_GROUP))
```
Note that the ranges of each age group are not consistent, which may lead to issues when interpreting the results.

```{r barplot sex ratio, echo=TRUE}
table(NYPD.2v$VIC_SEX)
barplot(table(NYPD.2v$VIC_SEX))
```
The gender proportion among the victims is extremely unbalanced. While this is a fact in itself, it may introduce some sort of bias during interpretation.

## 3-B) Plot; bivariate
```{r tabulate}
#tabulate
NYPD.2v <- as.data.frame(NYPD.2v)

# Create a cross table
NYPD.cross_table <- table(NYPD.2v$VIC_AGE_GROUP,NYPD.2v$VIC_SEX)

# Print the cross table
print(NYPD.cross_table)

barplot(NYPD.cross_table, beside=TRUE, legend = rownames(NYPD.cross_table), cex.axis = 0.5)
```
The difference in distribution doesn't seem apparent.
Let's try using regression analysis to get a formal support!

## 4-A) Formal test
```{r Chi-squared Goodness of Fit test}
age_female <- NYPD.cross_table[1:5,1]#exclude the unknown data
age_male   <- NYPD.cross_table[1:5,2]
chisq.test(age_female, age_male)
```

## 4-B) Binomial Regression Modeling: Sex~age

```{r discriminative_model_M/F}
# convert factor variable into numeric (M=1, F=0)
NYPD.2v <- NYPD.2v %>% mutate(SEX_num = case_when(
  VIC_SEX == "M" ~ 1,
  VIC_SEX == "F" ~ 0))
# View the data frame to confirm changes
#head(NYPD.2v)

# simple linear modeling
mod <- glm(formula = SEX_num~VIC_AGE_GROUP, data=NYPD.2v,
            family = binomial)
summary(mod)
```
All age factors are statistically **insignificant** in predicting the sex of victims, according to the dataset.
So, we cannot predict the victim's sex from their age information.

# Discussion on Bias
- This data is from New York. However, in certain regions, age information might be meaningful for making such predictions.

- The data focuses solely on individuals who have either committed crimes or become victims. Ignoring the information about people who are not included in the data could lead to misunderstandings.